The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-10-24 Paper.qmd
Major edits to paper to reflect switch to Lasso models from glmmTMB.

---
title: "The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study"
execute:
  echo: false
  warning: false
format:
  html:
    toc: true
lightbox: true
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
crossref:
  custom:
    - kind: float
      reference-prefix: Figure S
      key: sfig
      space-before-numbering: false
    - kind: float
      reference-prefix: Table S
      key: stbl
      space-before-numbering: false
---

@@

|  |  |
|------------------------------------|------------------------------------|
| Variable | Survey question wording |
| Crying Frequency | How often does your child cry? |
| Sad Frequency | How often is your child sad? |
| Tantrum Frequency | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency | How often does your child leave the house without telling you when they get mad? |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child? |
| Child Alloparenting Frequency | How often does your child take care of their other siblings? |

@@

We used 6-point Likert scales for parental reports of 1) the relative neediness of each participating child, and 2) the relative investment they provide to each participating child compared to the rest of their children. See Table @stbl-relvars for exact wording.

|  |  |
|------------------------------------|------------------------------------|
| Variable | Survey question wording |
| Relative Need | How much time, effort, and resources does this child need compared to your other children? |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

: Variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more. {#stbl-relvars}

@@

We asked caretakers about the last time each participating child was sad or cried. Caretakers provided a free response of the reason(s) for the emotional display, and their responses to it. The researcher then binned responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective). For the full list of options see Table S1. We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective). MG also coded the causes of each as follows: 1) the child was asked to do an *unwanted task*, 2) *punishment*, 3) *family conflict*, and 4) *discomfort, pain, injury, or illness*.

### Caretaker perceptions of child’s ability to avoid illness

@@

@@

The survey also included many demographic questions about the caretakers and their households. Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. These data allowed us to calculate the number of children (total, younger, and older) and adults within the household. We also imputed data for caregiver age by setting the missing value to the mean. For the data used in analyses, this amounted to imputing data for one mother which affected two children.

## Child anthropometrics

@@

All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables. Whereas confirmatory studies test prespecified hypotheses, exploratory analyses prioritize discovery: they risk overfitting the data, generating false positives, but can reveal unexpected yet potentially important relationships that warrant future investigation [@tukey1977].

@@

Child signaling frequencies and signaling costs, which were overdispersed count variables, were analyzed using generalized linear mixed models with a negative binomial error family, fit in R with the glmmTMB package [@brooks2017]. Unless otherwise noted, explanatory variables included child age, child sex,, child illness avoidance, caretaker education, caretaker partner status, household income, the number of other children in the household, the number of the adults in the household, the frequency of conflict between the primary caretaker and the child, and an interaction between the frequency of child alloparenting effort and child sex.. A household random effect was included in all models.

Parental responses to child signaling, their perceptions of child need, and their reported investment in the child were coded as three-level ordinal variables, either negative, neutral or positive; or less than other children, same as other children, or more than other children. We analyzed these outcomes using ordinal logistic regression, fit in R using the polr function in the MASS package [@venables2002], with clustered standard errors by household [@zeileis2002].

@@

kable(custom.summarize(modeldf, longform_dict[1:18])) %>% kable_styling()
```

@@

## Anthropometric predictors of child signaling

There were negative relationships between our Body Fat Index and signal cost (`{r} stats$mBFc$BodyFat$str`), crying frequency (`{r} stats$mBFcry$BodyFat$str`), and tantrum frequency (`{r} stats$mBFtantrum$BodyFat$str`). See @stbl-bodyfat-models and @fig-effect_bodyfat. However, in all models examining the effect of body fat on signaling, two children with Body Fat Index values greater than 6 and low signaling frequencies had a large effect on model predictions, and after removing them from the models, our body fat index was no longer a significant predictor.

Although interactions of Body Fat Index and sex were not significant, our models predicted a negative relationship between body fat and summed signal cost, crying frequency, and tantrum frequency for girls but no relationship for boys.

Height Z-scores, weight Z-scores, BMI Z-scores, and body fat percentage residuals were not significant predictors of signaling behavior. See xxxxSI TABLESxxxx for regression tables.

::: {#fig-effect_bodyfat}
```{r}
signaling_data_plot_bodyfat2
```

The effect of body fat (scaled composite measure of subscapular and tricep skinfold thicknesses) on A) the summed cost of signaling behavior, B) the frequency of crying, and C) the frequency of temper tantrums. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @stbl-bodyfat-models for the corresponding regression table.
:::

@@

\[add sad freq2\] Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children. Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor. The end result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.

There were significant positive relationships between perceived relative neediness within the family and both the summed frequency of child signaling and child age. See @fig-effect_relativeneed and @stbl-need-models. There was also a positive relationship between summed signal cost and perceived relative neediness within the family, controlling for the same variables; see @stbl-need-models and @sfig-effect-relativeneed.

# FIX: error ribbons cross 1 and 0

::: {#fig-effect_relativeneed}
```{r, fig.width= 10}
signaling_plot_need_freq
```

@@

Signals are capable of eliciting better treatment because the information they provide alters the behavior of signal targets in ways that benefit the signaler. For this reason, we tested the association between caretaker perceptions of relative child need within the household and the caretaker's relative investment within the household. Consistent with signaling theory, our model predicts a greater proportion of caretakers investing more in children who they deem needier and a greater proportion of caretakers investing less in children they deem less needy, controlling for the age of the child, and household demographics. See @fig-plot_relative_investment and @stbl-need-investment.

::: {#fig-plot_relative_investment}
```{r, fig.width= 8}
plot_relative_investment
```

@@

```{r}
# \[do we want to mentioned distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NegativeResponseBinary > 0)` negative response). However, they were also more likely to respond in a neutral or negative manner rather than a positive one when caretakers believed signaling came in response to the child being punished (`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).
```

@@

There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective. However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (any statistic report here?), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?). See @fig-effect_last_signal_response and @stbl-last-signal-model.

::: {#fig-effect_last_signal_response}
```{r, fig.width= 8}
signaling_plot_last_signal_response
```

@@

We found no significant associations between sex and child signaling behavior outside of the previously mentioned interactions, including when sex was included in models as interaction with age. See table SI xxxxxxx for regression tables which included the sex by age interaction. Monthly household income also was not a significant predictor of signaling behavior in any model nor was population when using language, immigration status, and neighborhood as proxies for it. See table SI xxxxxxx.

@@

We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys. In many societies, girls do more alloparenting than boys [@helfrecht2016 @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers [add info from an analysis here?]. If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.

@@

[delete based on lasso coefficients?]
However, child alloparenting effort was negatively associated with caregiver reports of child conflict. One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual. In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting. Ideally, we could control for signaling behavior to test this. However, the bidirectional relationship between child signaling and conflict with caretakers prevents this (explain better). It is also possible that our hypothesis about conflict over investment in the household is incorrect.

[delete now?]
*Body fat discussion*

Children with less body fat cried more, threw more temper tantrums, and had higher summed signal costs. In other words, needier children signaled more often. Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys. Female reproduction requires a significant caloric invest in children, both in the womb and when breastfeeding [@jasienska2009], costs which males do not have to pay. For this reason, low body fat, which can prevent ovulation in extreme cases [@ellison2003], is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.

@@

There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment. Caretakers also perceived older children as needier, despite having more conflicts with younger children. One reason for this could be that caretakers weighted monetary investment highly when evaluating child need.

Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households. However, crying frequency was associated with both height and weight gain for older children. If older children face greater reputation costs from crying, crying may be a relatively credible credible signal compared to crying among younger children, with it being a less viable strategy to exaggerate one's value of investment or the costliness of adversity.

@@

[delete?]
*In most models, there is a negative association between education and child signaling?*

@@

Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions. However, neighborhood quality was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients. One reason for this might that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness [no evidence of this when neighborhood quality is included in models predicting relative child need. ]

[old]
Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior. Although we sought data from both English-speaking participants born on the island and Spanish-speaking immigrants from the mainland, our smaller sample of English-speaking Utilians may have limited our ability to detect real differences between the populations. It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing different types adversity or weighting different types of investment differently. It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting. However, as previously mentioned, it is conceivable that household level differences have little effect on child signaling behavior.

@@

|  |  |
|------------------------------------|------------------------------------|
| Variable | Survey question wording |
| IllnessAvoidance1 | If an illness is 'going around', my child will get it. |
| IllnessAvoidance2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| IllnessAvoidance3 | My child’s immune system protects him/her from most illnesses that other kids get. |

@@

List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried.

@@

@@

## Regression tables

### Main models

::: {#stbl-main-models}
```{r}
modelsummary(list(
  Sadness = glmmTMBmodels$Model$SadFreqNillness, 
  Crying = glmmTMBmodels$Model$CryFreqNillness, 
  Tantrum = glmmTMBmodels$Model$TantrumFreqNillness,
  SignalFreq = glmmTMBmodels$Model$SignalFreqillness,
  SignalFreqMax = glmmTMBmodels$Model$SignalFreqMaxillness,
  SignalCost = glmmTMBmodels$Model$SignalCostillness),
  gof_map = "nobs", 
  coef_rename = shortform_dict, 
  stars = c('*' = .05, '**' = .01, '***' = .001), 
  coef_omit = "(Intercept)")
```

Regression tables for our main exploratory models.
:::

### Conflict frequency model

::: {#stbl-conflict-dep-var}
```{r}
modelsummary(glmmTMBmodels$Model$ConflictFreqN, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

Regression table for the model examining the frequency of conflict between child and caretaker as the outcome variable.
:::

### b

::: {#stbl-alloparenting-dep-var}
```{r}
modelsummary(malloparenting, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

Regression table for the model examining the frequency of child alloparenting effort as the dependent variable.
:::

### Models of signaling predicted by body fat index

::: {#stbl-bodyfat-models}
```{r}
modelsummary(list(
  Sadness = glmmTMBmodels$Model$SadFreqNbodyfat, 
  Crying = glmmTMBmodels$Model$CryFreqNbodyfat,
  Tantrum = glmmTMBmodels$Model$TantrumFreqNbodyfat,
  SignalFreq = glmmTMBmodels$Model$SignalFreqbodyfat,
  SignalFreqMax = glmmTMBmodels$Model$SignalFreqMaxbodyfat,
  SignalCost = glmmTMBmodels$Model$SignalCostbodyfat),
  gof_map = "nobs",
  coef_rename = shortform_dict,
  stars = c('*' = .05, '**' = .01, '***' = .001),
  coef_omit = "(Intercept)")
```

Regression tables for the models examining the effect of body fat on signaling.
:::

### d

::: {#stbl-need-models}
```{r}
modelsummary(list(SignalFreq = mN1, SignalCost = mN2), vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

Regression tables for the models examining relative perceptions of child need within the family as a dependent variable.
:::

:::

### Model of relative investment within the household

::: {#stbl-need-investment}
```{r}
modelsummary(polrModels$Model$Invest, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```
:::

### Model of caregiver response to last signal of need

::: {#stbl-last-signal-model}
```{r}
modelsummary(polrModels$Model$Response, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

Regression table for the model examining responses to the last time a child was sad or cried.
:::

### OLD Discussion

As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons. One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults. Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity of conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995]. For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\]. That said, all else may often not be equal if investment in older children is costlier, something which may be reflected in caretakers seeing older children as relatively needy. The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned. ~~It is also possible that parents see older children less than younger children. However, the negative association between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.~~

There was no significant association between the frequency of child sadness and age. Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011], from which, others can still obtain cue-based information about one's emotional state, well-being, and desires. By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals. For this reason, a change in signals that arise from sadness outside of crying (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often. It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.

>>>>

---
title: "The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study"
author: 
  - name: Michael R. Gaffney
    corresponding: false
    affiliation: 
      - name: Washington State University
        department: Anthropology
  - name: Edward H. Hagen
    orcid: 0000-0002-5758-9093
    email: edhagen@wsu.edu
    corresponding: true
    affiliation: 
      - name: Washington State University
        department: Anthropology
execute:
  echo: false
  warning: false
format:
  html:
    toc: true
    toc-expand: 2
    number-sections: true
lightbox: true
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
crossref:
  custom:
    - kind: float
      reference-prefix: Figure S
      key: sfig
      space-before-numbering: false
    - kind: float
      reference-prefix: Table S
      key: stbl
      space-before-numbering: false
---

@@

|  |  |
|--------------------|----------------------------------------------------|
| Variable | Survey question wording |
| Crying Frequency | How often does your child cry? |
| Sad Frequency | How often is your child sad? |
| Tantrum Frequency | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency | How often does your child leave the house without telling you when they get mad? |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child? |
| Child Alloparenting Frequency | How often does your child take care of their other siblings? |

@@

We used 6-point Likert scales for parental reports of 1) the relative neediness of children within the household, and 2) the relative investment they provide to each participating child compared to the rest of their children. Both questions were framed need as involving time, effort, and resources. See Table @stbl-relvars for exact wording.

@@

We asked caretakers about the last time each participating child was sad or cried. Caretakers provided a free response of the reason(s) for the emotional display, and their responses to it. The researcher then binned responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective). For the full list of options see @stbl-response-valence. We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective). MG and KS also coded the causes for the presence or absence of 17 elements (e.g., whether the caused involved the child was asked to do an *unwanted task*, *punishment*, *family conflict*, and/or *discomfort, pain, injury, or illness*).

### Caretaker perceptions of child health and ability to avoid illness

Caretakers were asked about the frequency of their child missing school and their child's history of medical problems in general terms (i.e., the extent to which they experienced a "variety of medical issues"). Both involved the use of five point Likert scales, the scores of which were averaged for use in analyses. See @stbl-medical-record for the wording of these variables.

@@

### Caretaker perceptions of house quality, neighborhood quality, home instability, and food security

Caregivers were asked various questions about their home, which we used to create a home quality index. This involved scoring the following summing the affrimatives to the following conditions: having an indoor bathroom, having a bathroom that is not shared with anyone outside of the household, having a flush toilet, having running water to wash hands in the bathroom, and having a home which did not flood in the last year.

Caregiver perceptions of neighborhood quality were measured using ten questions on a five point Likert scale relating to the costs and benefits associated with living in one's neighborhood. The mean of these values was used in analyses as a neighborhood quality index. See @stbl-Neighborhood-Quality for variable wording.

Levels of home instability were measured with four questions covering the degree to which the caregiver's family had/has to move homes and whether people move in and out of their homes frequently. All questions invovled five point Likert scales. See @stbl-Home-Instability for variable wording.

We used a modified version of the USDA Food Security Questionnaire [@blumberg1999] to measure food insecurity, with the most significant changes being converting of the measures to a five point Likert scale and averaging them for a food insecurity index. See @stbl-Food-Security for the wording of these variables.

@@

The survey also included many demographic questions about the caretakers and their households. Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) kinship ties within households, (4) partnership status of the primary caregiver, (5) number of adults who provide childcare within the household, (6) highest level of education of the primary caregiver, (7) immigration status (in regards to Utila), (8) monthly household income, (9) neighborhood of residence, (10) current job of the primary caretaker, and (11) ownership of various possesions (e.g., homes, washing machines, or cellphones; see SIXXXXXX for full list). These data allowed us to calculate the number of children (total, younger, and older) and adults within the household as well as the genetic relatedness between siblings. We also imputed data for caregiver age by setting the missing value to the mean. For the data used in analyses, this amounted to imputing data for one mother which affected two children.

\[mention Campanado here and in ethnographic portion?\]

\[I think this gets deleted or do we mention no findings worth reporting?\] \## Child anthropometrics

@@

All statistical analyses were exploratory. Whereas confirmatory studies test prespecified hypotheses, exploratory analyses prioritize discovery: they risk overfitting the data, generating false positives, but can reveal unexpected yet potentially important relationships that warrant future investigation [@tukey1977].

@@

To investigate the associations between potential causes of child signaling and conflict with caretakers and caregiver reports of child behavior, we used elasticnet regression models. Elasticnet models (e.g., lasso and ridge regression) are well suited to instances when the number of predictor variables are large relative to the number of observations due their use of penalty terms and cross-validation to limit the risk of overfitting the data.

Analyses were run in R using the glmnet package [@friedman2010] with a quasipoisson error family due to our outcome variables being overdispered counts. Predictors includeddemographic, household, health, and neighborhood variables that might cause signaling behavior or conflict between children and caretakers. All variables were centered and standardized by one standard deviation prior to fitting.

Parental responses to child signaling, their perceptions of child need, and their reported investment in the child were coded as three-level ordinal variables, either negative, neutral or positive; or less than other children, same as other children, or more than other children. We analyzed these outcomes using ordinal logistic regression with elastic net penalties, fit in R using the ordinalNet package [@wurm].

@@

kable(custom.summarize(modeldf, longform_dict[1:18])) %>% kable_styling()
# dx <- modeldf
# names(dx) <- shortform_dict[names(modeldf)] # set-names??? line 1844
# datasummary_skim(dx, type = "categorical)
```

@@

@@

Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children. Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor. The end result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.

There were positive relationships between perceived relative neediness within the family and both child age and the frequency of child sadness. See @fig-effect_relativeneed.

::: {#fig-effect_relativeneed}
![](Figures/plot_need_combined.svg)

::: {#fig-effect_relativeneed2}
```{r, fig.width= 10}
signaling_plot_need_freq
# plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "A")
```

@@

Signals are capable of eliciting better treatment because the information they provide alters the behavior of signal targets in ways that benefit the signaler. For this reason, we tested the association between caretaker perceptions of relative child need within the household and the caretaker's relative investment within the household. Consistent with signaling theory, our model predicts a greater proportion of caretakers investing more in children who they deem needier and a greater proportion of caretakers investing less in children they deem less needy, controlling for the age of the child, and household demographics. See @fig-plot_relative_investment.

::: {#fig-plot_relative_investment}
![](Figures/plot_invest_need.svg)

::: {#fig-plot_relative_investment2}
```{r, fig.width= 8}
plot_relative_investment
#plot_invest_need
```

@@

```{r}
# \[do we want to mention distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NegativeResponseBinary > 0)` negative response). However, they were also more likely to respond in a neutral or negative manner rather than a positive one when caretakers believed signaling came in response to the child being punished (`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).
```

@@

Our models found no evidence of relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective. However, primary caretakers were  more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?). See @fig-effect_last_signal_response.

::: {#fig-effect_last_signal_response}
![](Figures/plot_caregiver_response_combined.svg)

::: {#fig-effect_last_signal_response2}
```{r, fig.width= 8}
signaling_plot_last_signal_response
# plot_caregiver_punish / plot_caregiver_pain + ggtitle("") + plot_layout(guides = 'collect')
```

@@

[needs updating. worth having a section just for this?]
We found no evidence of associations between sex and child signaling behavior outside of the previously mentioned interactions. Monthly household income also was not a significant predictor of signaling behavior in any model nor was population when using language, immigration status, and neighborhood as proxies for it.

@@

We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys. In many societies, girls do more alloparenting than boys [@helfrecht2016 @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers \[add info from an analysis here?\]. If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.

@@

\[delete based on lasso coefficients?\] However, child alloparenting effort was negatively associated with caregiver reports of child conflict. One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual. In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting. Ideally, we could control for signaling behavior to test this. However, the bidirectional relationship between child signaling and conflict with caretakers prevents this (explain better). It is also possible that our hypothesis about conflict over investment in the household is incorrect.

@@

There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment.

Caretakers perceived older children as needier, despite having more conflicts with younger children. One possible reason for this might be that caretakers see the types of investment older children seek as more costly. In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often during surveys. It may also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.

[still needs updating]
Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households. However, crying frequency was associated with both height and weight gain for older children. If older children face greater reputation costs from crying, crying may be a relatively credible credible signal compared to crying among younger children, with it being a less viable strategy to exaggerate one's value of investment or the costliness of adversity.

@@

\[delete?\] *In most models, there is a negative association between education and child signaling?*

@@

Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions. However, neighborhood quality was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients. One reason for this might that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness \[no evidence of this when neighborhood quality is included in models predicting relative child need. \]

\[old\] Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior. Although we sought data from both English-speaking participants born on the island and Spanish-speaking immigrants from the mainland, our smaller sample of English-speaking Utilians may have limited our ability to detect real differences between the populations. It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing different types adversity or weighting different types of investment differently. It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting. However, as previously mentioned, it is conceivable that household level differences have little effect on child signaling behavior.

@@

## Additional information for survey measures

::: {#stbl-relvars}

|  |  |
|----------------------|--------------------------------------------------|
| Variable | Survey question wording |
| Relative Need | How much time, effort, and resources does this child need compared to your other children? |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

Variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more. 
:::

|  |  |
|------------------|------------------------------------------------------|
| Variable | Survey question wording |
| SchoolAges1 | My child has missed a lot of school due to illness. |
| SchoolAges2 | My child has had to deal with a variety of medical issues in their life so far. |

Variable names and question wording for measures about a child's health history. Both questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." {#stbl-medical-record}

|  |  |
|-----------------------|-------------------------------------------------|
| Variable | Survey question wording |
| IllnessAvoidance1 | If an illness is 'going around', my child will get it. |
| IllnessAvoidance2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| IllnessAvoidance3 | My child’s immune system protects him/her from most illnesses that other kids get. |

@@

|  |  |
|------------------|------------------------------------------------------|
| Variable | Survey question wording |
| FoodSecurity_1 | In the past year, the food that we bought often didn’t last, and we didn’t have money to get more. |
| FoodSecurity_2 | In the past year, we couldn’t afford to eat balanced meals. |
| FoodSecurity_3 | In the past year, I (or other adults in my household) sometimes cut the size of my meals or skipped meals because there wasn't enough money for food. |
| FoodSecurity_4 | In the past year, I sometimes ate less than I felt I should because there wasn't enough money for food. |
| FoodSecurity_5 | In the past year, I was sometimes hungry but didn’t eat because there wasn’t enough money for food. |
| FoodSecurity_6 | In the past year, I ate lower quality food because I couldn't afford better food. |

: Variable names and question wording for the modified version of the USDA Food Security Questionnaire [@blumberg1999]. All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.” {#stbl-Food-Security}

|  |  |
|--------------------|----------------------------------------------------|
| Variable | Survey question wording |
| HomeInstability_1 | We move our home a lot. |
| HomeInstability_2 | I wish our household felt more stable. |
| HomeInstability_3 | I worry that we will be forced to move before we are ready. |
| HomeInstability_4 | People move in and move out of our home frequently. |

: Variable names and question wording for the home instability questions. All questions were asked using a Likert scale from 1-5 where 1 was labeled “Disagree strongly” and 5 was labeled “Agree Strongly.” {#stbl-Home-Instability}

|  |  |
|-----------------------|-------------------------------------------------|
| Variable | Survey question wording |
| NeighborhoodQuality_1 | As a place to live, my neighborhood is good quality. |
| NeighborhoodQuality_2 | As a place to raise children, my neighborhood is good quality. |
| NeighborhoodQuality_3 | My neighborhood quality has gotten worse over time. |
| NeighborhoodQuality_4 | My neighborhood is more dangerous than other parts of the island. |
| NeighborhoodQuality_5 | Sometimes I hear about a fight with a weapon in my neighborhood. |
| NeighborhoodQuality_6 | Sometimes I hear about illegal drug activity in my neighborhood. |
| NeighborhoodQuality_7 | Sometimes I hear about someone being badly hurt in my neighborhood. |
| NeighborhoodQuality_8 | Sometimes I hear about someone's home being robbed in my neighborhood. |
| NeighborhoodQuality_9 | Sometimes I hear about someone being mugged in my neighborhood. |
| NeighborhoodQuality_10 | Sometimes I hear about someone being assaulted in my neighborhood. |

: Variable names and question wording for the perceptions of neighborhood quality questions. All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.” {#stbl-Neighborhood-Quality}

@@

: List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried. {#stbl-response-valence}

List of possessions owned (yes or no).

Motorized vehicle (scooter, 4-wheeler, etc.) Bicylce Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card

@@
